\section{Individual Part: Birth-and-death vectors (Clemente Concha Kostin)}
\subsubsection*{Introduction and Motivation}
This model analyzes vector population dynamics ($S_v, I_v$), focusing on birth ($B_v$), death ($D_v$), and infection processes. Human infectious pressure is assumed constant to isolate how these demographics impact vector infection status.

\subsection*{Assumptions}

\begin{itemize}[nosep,leftmargin=*,topsep=0.5ex] 
    \item Vectors are always born susceptible.
    \item Vectors do not recover and remain infectious for life.
    \item Vector birth rate is proportional to the total current vector population.
    \item Vectors die at a constant per capita natural death rate.
    \item The force of infection on vectors (per capita rate of new vector infections) is constant.
\end{itemize}

\subsection*{Model Variables}
\noindent
\begin{tabularx}{\linewidth}{@{}l X@{}} % 
\toprule
\textbf{Variable} & \textbf{Description} \\
\midrule
$S_v(t)$ & Number of susceptible vectors at time $t$. \\
$I_v(t)$ & Number of infectious vectors at time $t$. \\
$N_v(t)$ & Total vector population, $N_v(t) = S_v(t) + I_v(t)$. \\
\bottomrule
\end{tabularx}

\subsection*{Model Parameters}
\noindent 
\begin{tabularx}{\linewidth}{@{}l X@{}}
\toprule
\textbf{Parameter} & \textbf{Description} \\
\midrule
$B_v$       & Per capita birth rate of vectors (total new vectors: $B_v N_v$). \\
$D_v$       & Per capita natural death rate of vectors. \\
$\lambda_v^*$ & Constant force of infection on vectors (per capita rate). \\
\bottomrule
\end{tabularx}

\subsection*{System of Differential Equations}
The dynamics of the vector population are described by:
\begin{align} 
\frac{dS_v}{dt} &= B_v N_v - \lambda_v^* S_v - D_v S_v \label{eq:clemente_dSv_simplified} \\
\frac{dI_v}{dt} &= \lambda_v^* S_v - D_v I_v \label{eq:clemente_dIv_simplified}
\end{align}
New susceptible vectors are born at rate $B_v N_v$. Susceptible vectors get infected at rate $\lambda_v^* S_v$. Both $S_v$ and $I_v$ die at per capita rate $D_v$.
\\\\
The total vector population $N_v(t)$ changes according to:
\begin{align*} % Using align* for unnumbered derivation
\frac{dN_v}{dt} &= \frac{dS_v}{dt} + \frac{dI_v}{dt} \\
               &= (B_v N_v - \lambda_v^* S_v - D_v S_v) + (\lambda_v^* S_v - D_v I_v) \\
               &= B_v N_v - D_v S_v - D_v I_v \\
               &= B_v N_v - D_v (S_v + I_v) \\
               &= B_v N_v - D_v N_v = (B_v - D_v)N_v
\end{align*}
This core equation, $\frac{dN_v}{dt} = (B_v - D_v)N_v$, shows total vector population size is driven by its per capita birth ($B_v$) and death ($D_v$) rates. These demographics, with infection rate $\lambda_v^*$, determine the number of infectious vectors.

% --- End of content for Chapters/3. Individual Part (Clemente).tex ---