\section{Individual part: Birth-and-rate humans (Tim van Wijk)}

\subsection{Motivation}
In the standard SI-SI model, many assumptions are made that are not realistic. I will propose a modified version of the SI-SI model taking into account the birth-and-death rate of humans. This modified model will be slightly more realistic compared to the original SI-SI model.

\subsection{Assumptions}
See 1.1.1
Extra:
For the proposed model, we assume that babies are born without disease and that both infected and susceptible people can die. To simulate the birth-and-death rate of humans in the model, we will denote the number of births on day $t$ by $B(t)$. We will split the number of deaths into deaths of people who are infected on day $t, D_I(t)$ and those who are susceptible on day $t, D_S(t)$. All three of these functions are a function of time and affect the number of total hosts. We get $N_H(t)=S_H(t) + I_H(t) + B(t)-D_I(t) -D_S(t)$ for the number of humans. The number of vectors is not affected by the birth-and-death rate of humans and therefore, $N_V = S_V(t) + I_V(t)$. Let $\gamma N_H(t)$ equal the number of bites of vectors on day $t$ per vector. Then the total number of bites of vectors on day $t$ is $N_V \gamma N_H(t)$. We let $P_{V\rightarrow H}$ be the probability that a single bite from an infected vector infects a susceptible human and $P_{H\rightarrow V}$ the probability that a susceptible vector gets infected by a bite from an infected human. The total number of times a person is bitten can be expressed as $N_V \gamma$.  We let $\beta_H = \gamma N_V P_{V\rightarrow H}$. We can express the total number of new infected hosts per time unit as $\beta_H \frac{I_V}{N_V} S_H$. We do the same for vectors. Let $\beta_V = \gamma N_H(t) P_{H\rightarrow V}$. Then the total number of new infected vectors per time unit is $\beta_V \frac{I_H}{N_H(t)} S_V$. This leads to the following system of ODEs:
\begin{equation}
S_H'(t) = B(t) -\beta_H \frac{I_V}{N_V} S_H - D_S(t) 
\end{equation}
\begin{equation}
I_H'(t) = \beta_H \frac{I_V}{N_V} S_H - D_I(t)
\end{equation}
\begin{equation}
S_V' = -\beta_V \frac{I_H}{N_H(t)} S_V
\end{equation}
\begin{equation}
I_V' = \beta_V \frac{I_H}{N_H(t)} S_V
\end{equation}